Carp\'{e} Data: Supporting Serendipitous Data Integration in Personal Information Management

### 2012-09-19 chi2012.tex
Ran the entire thing through ispell with a british englosh dictionary.

% Load basic packages
\usepackage{balance}  % to better equalize the last page
\usepackage{graphics} % for EPS, load graphicx instead
\usepackage{times}    % comment if you want LaTeX's default font
\usepackage{url}      % llt: nicely formatted URLs
\usepackage{multirow} % multirow tables

@@

In recent years, an unprecedented quantity and variety of information has been made available as structured data on the Web.  This information has emerged from many channels, from the APIs now offered by most of the major retailers, apps, and social media sites, to the massive publishing of datasets by governmental and other institutions, to even the emergence of new kinds of data sources, such as wearable and smartphone-based biosensors.  A primary goal of making much of this data available has been to give end-user citizens the ability to make more informed decisions pertaining to their health, wealth, and well-being \cite{Shadbolt:2006:SWR:1155313.1155373}.  This data is predominately used by specialised app developers, journalists and other ``data specialists'' but remains inaccessible to citizens because of the available tools.  For example, the most commonly used structured personal information management (PIM) tools either manage a small, fixed set of data types (such as digital calendaring tools and  to-do list managers) or provide little or no support for structured data at all, such as text editors and sketching/drawing tools \cite{infoscraps}.

@@

% i think we do not need to differentiate between semantic web and databse integration here, we're saying they're both bad, let's lump them together to keep the argument flowing and not kill our readers
The database integration and Semantic Web \cite{Shadbolt:2006:SWR:1155313.1155373} research communities have primarily focused efforts on purely-automatic approaches to the problem of data integration.  Such approaches typically take the form of either \emph{ontology} or \emph{schema-level matching}, or \emph{direct} or \emph{instance-level} matching.  We presently focus  on \emph{instance-level} matching, as schemas may not always be available, nor easily accesssed by users. Approaches to automatic instance matching have included applied natural language processing (NLP) techniques for terminological matching (e.g., \cite{euzenat2004api}), machine-learning techniques that use both structural and terminological features for matching \cite{doan2003learning}, probabilistic approaches \cite{suchanek2011paris}, and instance schema comparisons using structural and logical definitions \cite{castano2006matching}.   In real-world settings, however, manual and semi-automatic matching approaches are favoured over automatic because of the need for high degrees of accuracy.

@@

\section{Pre-studies: The Need for and Challenges in realizing Ad-Hoc Data Integration in PIM Tools}

@@

\subsubsection{Enumerated-type value consolidation}
PS.2 revealed that enumerated values of properties were often inconsistent, e.g. ``Casual dining'' vs ``Relaxed'' for a restaurant's ``atmosphere'' property. To reconcile and consolidate corresponding values from different  datasources, we extended the drag and drop support of co-reference combination to also allow users to drag and drop to combine enumerated literal values as well.

@@

%% it would be nice if we had space to summarize but i fear we don't

%% \subsection{Summary of interaction capabilities}
%% \begin{itemize}
%% \item \emph{Group multi-path value selection} - Selecting any of the %% properties of an instance in a group (window) causes all of the rest %% of the instances that have that property in that group to be %% dereferenced along the same property, displaying the resulting value %% underneath the instance label.  Selecting a further property of the %% resulting value further dereferences this value.
%% \item \emph{Drag-and-drop coreference / value combination} - In order %% to specify that two instances correspond to the same entity, they can %% be simly dragged together
%% \item \emph{Group-based Drag-and-drop matching} - In order to specify %% that two instances correspond to the same entity, they can be simly %% dragged together
%% \item \emph{Universal model-based brushing} -
%% \end{itemize}

@@

% heather: instead of these low-level hypotheses, i've made higher level ones that we can explicitly address in the disucssion section - the more hypotheses we have, the more we have to come back to them; since we're short on space we might as well keep it simple!

@@

\subsection{Experimental Setup}
The participants were given a standard desktop PC running MacOS X on a 22-inch monitor with a standard keyboard and mouse.  During the studies the Excel workbook and DP were initially set to full screen.  Each participant was asked to complete two tasks, one A and one B condition.  Each participant was given a maximum of 10 minutes to complete a task, but were told that if they finished earlier they could tell us and move on to the next condition.  In order to select which tasks the participant was given, we used a latin square to permutate the tasks and conditions.  In order to train the user on how to use our interface, they viewed a five-minute DataPalette tutorial video introducing its basic functionality, how to use histograms and maps, and linking instances.

@@

%Each study consisted of two tasks, one A and one B condition.  Each participant was given 10 minutes to complete a task.  In order to select which tasks the participant was given, we used a latin square to permutate the tasks and conditions.  In order to train the user on how to use our interface, they viewed a five-minute DataPalette tutorial video introducing its basic functionality, how to use histograms and maps, and linking instances.

@@

Twenty-six (26) participants responded, from which we selected the first 20.  Ten participants were female.  All of the participants spoke English, but 12 of them were non-native English speakers.  Seven were non-students, comprising University staff and alumni; out of the remaining 13, eight studied computer science and 5 studied other subjects.   On average, each study took 40 mins to complete.  The information in Table \ref{tbl:data} summarises the metrics that were captured of the trials for each participant. We analyse this data next.

@@

%What we found 
%
%What we didn't find (evidence for)
%
%What we might look at next -- follow-on work
%		We focused on terminological heterogeneity -- didn't address structural heterogeneity, semantic heterogeneity
%		We focused on finite (small) sets Streaming data (how do select future data --) 
%		Scale - small data, what are the facilites required for large data
%		Collaborative - how might interaction support multiple users collaborating?
%
%Limitations found from study... what is the take away from the limitations

@@

%% In this paper, we take the position that there is a need for a difference in focus between the approaches taken to data integration between the approaches taken in database integration and that of PIM, because of vast differences in requirements.  In database systems, the overall objective is usually to allow multiple large databases to be accessed as if one. This requires essentially full mapping of all data types and objects and well-defined semantics concerning where and how data gets converted. The upfront costs of such complete integration are typically high, but are typically justified in the long-term, high-volume use cases imagined. In typical sensemaking scenarios of PIM, however, people might need to combine information from multiple sources for a one-off task; meaning that any upfront cost would directly impact the time and effort a person has allocated for their task.  Moreover, while a fully, uniform query capability might be nice, such capability is likely of less overall importance for a single quick task than simply being able to quickly and easily identify and cope with differences.

>>>>

% Load basic packages
\usepackage{balance}  % to better equalise the last page
\usepackage{graphics} % for EPS, load graphicx instead
\usepackage{times}    % comment if you want LaTeX's default font
\usepackage{url}      % llt: nicely formatted URLs
\usepackage{multirow} % multirow tables

@@

In recent years, an unprecedented quantity and variety of information has been made available as structured data on the Web.  This information has emerged from many channels, from the APIs now offered by most of the major retailers, apps, and social media sites, to the massive publishing of datasets by governmental and other institutions, to even the emergence of new kinds of data sources, such as wearable and smartphone-based bio-sensors.  A primary goal of making much of this data available has been to give end-user citizens the ability to make more informed decisions pertaining to their health, wealth, and well-being \cite{Shadbolt:2006:SWR:1155313.1155373}.  This data is predominately used by specialised app developers, journalists and other ``data specialists'' but remains inaccessible to citizens because of the available tools.  For example, the most commonly used structured personal information management (PIM) tools either manage a small, fixed set of data types (such as digital calendaring tools and  to-do list managers) or provide little or no support for structured data at all, such as text editors and sketching/drawing tools \cite{infoscraps}.

@@

% i think we do not need to differentiate between semantic web and database integration here, we're saying they're both bad, let's lump them together to keep the argument flowing and not kill our readers
The database integration and Semantic Web \cite{Shadbolt:2006:SWR:1155313.1155373} research communities have primarily focused efforts on purely-automatic approaches to the problem of data integration.  Such approaches typically take the form of either \emph{ontology} or \emph{schema-level matching}, or \emph{direct} or \emph{instance-level} matching.  We presently focus  on \emph{instance-level} matching, as schemas may not always be available, nor easily accessed by users. Approaches to automatic instance matching have included applied natural language processing (NLP) techniques for terminological matching (e.g., \cite{euzenat2004api}), machine-learning techniques that use both structural and terminological features for matching \cite{doan2003learning}, probabilistic approaches \cite{suchanek2011paris}, and instance schema comparisons using structural and logical definitions \cite{castano2006matching}.   In real-world settings, however, manual and semi-automatic matching approaches are favoured over automatic because of the need for high degrees of accuracy.

@@

\section{Pre-studies: The Need for and Challenges in realising Ad-Hoc Data Integration in PIM Tools}

@@

\subsubsection{Enumerated-type value consolidation}
PS.2 revealed that enumerated values of properties were often inconsistent, e.g. ``Casual dining'' vs ``Relaxed'' for a restaurant's ``atmosphere'' property. To reconcile and consolidate corresponding values from different  data sources, we extended the drag and drop support of co-reference combination to also allow users to drag and drop to combine enumerated literal values as well.

@@

%% it would be nice if we had space to summarise but i fear we don't

%% \subsection{Summary of interaction capabilities}
%% \begin{itemize}
%% \item \emph{Group multi-path value selection} - Selecting any of the %% properties of an instance in a group (window) causes all of the rest %% of the instances that have that property in that group to be %% dereferenced along the same property, displaying the resulting value %% underneath the instance label.  Selecting a further property of the %% resulting value further dereferences this value.
%% \item \emph{Drag-and-drop coreference / value combination} - In order %% to specify that two instances correspond to the same entity, they can %% be simply dragged together
%% \item \emph{Group-based Drag-and-drop matching} - In order to specify %% that two instances correspond to the same entity, they can be simply %% dragged together
%% \item \emph{Universal model-based brushing} -
%% \end{itemize}

@@

% heather: instead of these low-level hypotheses, I've made higher level ones that we can explicitly address in the discussion section - the more hypotheses we have, the more we have to come back to them; since we're short on space we might as well keep it simple!

@@

\subsection{Experimental Setup}
The participants were given a standard desktop PC running MacOS X on a 22-inch monitor with a standard keyboard and mouse.  During the studies the Excel workbook and DP were initially set to full screen.  Each participant was asked to complete two tasks, one A and one B condition.  Each participant was given a maximum of 10 minutes to complete a task, but were told that if they finished earlier they could tell us and move on to the next condition.  In order to select which tasks the participant was given, we used a Latin square to permutate the tasks and conditions.  In order to train the user on how to use our interface, they viewed a five-minute DataPalette tutorial video introducing its basic functionality, how to use histograms and maps, and linking instances.

@@

%Each study consisted of two tasks, one A and one B condition.  Each participant was given 10 minutes to complete a task.  In order to select which tasks the participant was given, we used a Latin square to permutate the tasks and conditions.  In order to train the user on how to use our interface, they viewed a five-minute DataPalette tutorial video introducing its basic functionality, how to use histograms and maps, and linking instances.

@@

Twenty-six (26) participants responded, from which we selected the first 20.  Ten participants were female.  All of the participants spoke English, but 12 of them were non-native English speakers.  Seven were non-students, comprising University staff and alumni; out of the remaining 13, eight studied computer science and 5 studied other subjects.   On average, each study took 40 minutes to complete.  The information in Table \ref{tbl:data} summarises the metrics that were captured of the trials for each participant. We analyse this data next.

@@

%What we found 
%
%What we didn't find (evidence for)
%
%What we might look at next -- follow-on work
%		We focused on terminological heterogeneity -- didn't address structural heterogeneity, semantic heterogeneity
%		We focused on finite (small) sets Streaming data (how do select future data --) 
%		Scale - small data, what are the facilities required for large data
%		Collaborative - how might interaction support multiple users collaborating?
%
%Limitations found from study... what is the take away from the limitations

@@

%% In this paper, we take the position that there is a need for a difference in focus between the approaches taken to data integration between the approaches taken in database integration and that of PIM, because of vast differences in requirements.  In database systems, the overall objective is usually to allow multiple large databases to be accessed as if one. This requires essentially full mapping of all data types and objects and well-defined semantics concerning where and how data gets converted. The upfront costs of such complete integration are typically high, but are typically justified in the long-term, high-volume use cases imagined. In typical sense-making scenarios of PIM, however, people might need to combine information from multiple sources for a one-off task; meaning that any upfront cost would directly impact the time and effort a person has allocated for their task.  Moreover, while a fully, uniform query capability might be nice, such capability is likely of less overall importance for a single quick task than simply being able to quickly and easily identify and cope with differences.
